% Phase-plane analysis using the *phaseR* package
% Stephen Eglen
% 2016-10-25

## Phase plane analysis of Sheep--Rabbit system

Here we will look in detail at the rabbit-sheep system (Strogatz,
p155).  We first load the packages that we need:

```{r}
require(phaseR)
require(deSolve)
```

The state of the system is a two-dimensional vector, v=(r,s), where r
is the rabbit population, and s is the sheep population.  There are no
extra parameters to pass to this system.

```{r}
Sheep=function(t,y,parameters) {
  r=y[1]; s=y[2]
  drdt = r*(3 - r - (2*s))
  dsdt = s * (2 - r - s)
  list(c(drdt, dsdt))
}
```

### Evolution from one initial condition

First we simulate the evolution of the system from one initial
condition, and find that in this case, the Rabbits win, and the sheep
lose.

```{r}
init.cond = c(1.5, 1.3)
times=seq(from=0, to=20, length=100)
out = lsoda(init.cond, times, Sheep)
matplot(out[,1], out[,-1], type='l',xlab='time',
         ylab='y1,y2',lwd=2)
legend('topleft', c('Rabbits', 'Sheep'), lty=c(1,2), col=c('black', 'red'))
```

## Phase plane analysis

Now we do the phase plane analysis.  In the following, y0 is a matrix
of different initial conditions to try out.

```{r}
flow <- flowField(Sheep, x.lim=c(0,3), y.lim=c(0,3), points=19,
                  xlab='Rabbits', ylab='Sheep', add=F)
grid()
nullclines <- nullclines(Sheep, x.lim=c(-1, 3.5), y.lim=c(-1,3.5), points=500)

y0 <- matrix( c(0.5, 0.5,  2.5, 1.0, 0.5, 1.5,  2.5, 2.5), 4, 2, TRUE)
traj <- trajectory(Sheep, y0=y0, t.end=10)
```

From the nullclines, we can read off the four steady-state points
which we hard code into ss matrix (one steady-state per row).  We add
an extra function mystablility to examine the classification of the
steady sate, to return a suitable point style.


```{r}
mystability = function(...) {
  st = stability(...)$classification
  res = switch(st,
               "Unstable node"=1,
               "Stable node"=19,
               "Saddle"=13)
  res
}

ss = matrix( c(0,0,  0,2,   3,0,  1,1),ncol=2, byrow=T)
stab = apply(ss, 1, function(s) {mystability(Sheep, y.star=s)})
flow <- flowField(Sheep, x.lim=c(0,3), y.lim=c(0,3), points=19,
                  xlab='Rabbits', ylab='Sheep', add=F,
                  asp=1,
                  las=1)
nullclines <- nullclines(Sheep, x.lim=c(-1, 3.5), y.lim=c(-1,3.5), points=500)
points(ss[,1], ss[,2], pch=stab,cex=3)
```

(NB: in the above chunk, we repeat the calculation of the flow field
and the nullclines -- this is inefficient, but I see no immediate way
around it.)

## Compilation notes

```{r eval=FALSE}
rmarkdown::render('sheep-phaseR.Rmd')
```

